% !TEX root = 00_thesis.tex

\chapter[Acknowledgments]{Acknowledgments}

\squarepar{%

First and foremost, I want to thank Prof.\,Thiele for giving me the opportunity to work on my doctoral dissertation under his supervision: thank you for giving me a chance even though I came from a different field and background, I hope you do not regret it. I sincerely appreciated your dedication to good work, your availability, and our many heated discussions. I am glad I could sometimes convince you while we initially disagreed! Not often, but sometimes :-)

Next, I want to thank Thiemo and Martina for serving in my examination committee, having reviewed my dissertation, and the interesting discussions we had before and after. I am glad (and proud) that researchers whom I respect both the work and the personality recommended my dissertation for graduation.

Other these four years, I have been lucky enough to collaborate with many people from various groups and institutions: Licong, Carlo, Markus, Fabian, Dominic, Sebastian, Ramona, Usman, Laurent; thank you all very much for your support.
In particular, I want to thank Reto: your expertise as engineer and embedded system programmer has been invaluable for my work; you played a big role in enabling me to turn theory into practice, to implement and run my ideas. This was a lot of fun for me, I hope my endless questions were not too annoying for you :-)
Another set of great collaborators were the master students I had the pleasure to supervise: Jonas, Fabian, Alex, Andreas, Antonios, Jonathan, Jan, Anna, and Raphael. You all did terrific work! I hope you enjoyed your projects, our discussions, and the presentation rehearsal sessions as much as I did.

I also want to thank the mentors and supervisors I had before my doctorate. \linebreak
I would not have been able to do all this without the great training you provided during my studies at ENS Cachan (now ENS Paris-Saclay):
Jean-Jacques, Gregory, Jean-Marc, and Jean-Marie. My many thanks as well to Prof.\,Hedrick who offered my first real research opportunity in his group in US Berkeley.

Of course, my deepest gratitude goes to Marco: you offered countless hours of your time to our projects, discussions, and late-night paper writing sessions. You have been a great mentor and collaborator; I really learned tons from you. I hope life will let us find ways to keep working together; it was just so great!

\pagebreak

% Personal
On a more personal level, I would like to thank the whole TEC group for the fruitful time (and the not-so-fruitful-yet-funny-time) we spent together: Tonio, Reto, Jan, Georgia, Pengcheng, Roman, Felix, Andres, Balz, Olga, Rehan, Zimu, Philipp, Lukas, Stefan, Roman, Xiaoxi, Yun, Zhongnan, Naomi, and Andreas.
Naturally, the amazing TEC staff also plays a big role in making TEC a nice place to work at; many thanks to Rike, Susann, and of course Beat.
A special thanks to Matthias: you were always up for a debate on science, openness, politics, education, arts, and more; it was really nice to have someone I could share my fantasies with -- let's call that meaningful procrastination.

Naturally, a doctorate also demands support outside of the workplace. In this regard, my parents and family have always been amazing, physically far, but close to the heart. Many thanks also to my friends from the ASVZ Running and the Akademisher Alpine Clüb Zürich, who pushed me to discover and enjoy the wonders of outdoor activities in the Swiss Alps; I fear I am now addicted.

And of course, last but certainly not least, my infinite gratitude to my wife: your love and smile are indispensable to my life; you balance me out, share my joys, and help me becoming a better me. And you bake, so really, I am fulfilled.

}
