% !TEX root = 00_thesis.tex

\newcommand{\aut}[1]{\vspace*{0,5em}\\\noindent\newblock{#1}}
\newcommand{\tit}[1]{\newblock{\textbf{#1}}}
\newcommand{\con}[1]{\newblock{\emph{#1}}}
\newcommand{\jou}[1]{\newblock{\emph{#1}}}
%\newcommand{\jou}[1]{\newblock{\emph{#1}}}
\newcommand{\loc}[1]{\newblock{#1}}
\newcommand{\rem}[1]{\newblock{#1}}

\chapter{List of Publications}

\begin{footnotesize}

The following list includes publications that form the basis of this thesis. The corresponding chapters are indicated in parentheses.
%\vspace{1.5cm}
\aut{Gomez A, Sigrist L, Magno M, Benini L, Thiele L.}
\tit{Dynamic Energy Burst Scaling for Transiently Powered Systems.}
\con{In Proceedings of Conference on Design, Automation \& Test in Europe (DATE).}
\loc{EDA Consortium. 2016.}
\rem{(\chref{ch:debs})}
%
%\\%
%\vspace{1.5cm}
\aut{Gomez A, Sigrist L, Schalch T, Benini L, Thiele L.}
\tit{Wearable, Energy-Opportunistic Vision Sensing for Walking Speed Estimation.}
\con{In Proceedings of Sensors Applications Symposium (SAS).}
\loc{IEEE. 2017.}
\rem{(\chref{ch:debs})}
%
\aut{Gomez A, Benini L, Thiele L. }
\tit{Designing the Batteryless IoT. }
\con{Design, Automation \& Test in Europe (DATE) PhD Forum.}
\loc{EDA Consortium 2017.}
\rem{(\chref{ch:debs})}
%\\%
%\vspace{1.5cm}
\aut{Gomez A, Sigrist L, Schalch T, Benini L, Thiele L.}
\tit{Efficient, Long-Term Logging of Rich Data Sensors using Transient Sensor Nodes.}
\jou{Transactions on Embedded Computing Systems.}
\loc{ACM. 2017.}
\rem{(\chref{ch:nvmh})}
%
\aut{Gomez A, Benini L, Thiele L. }
\tit{Designing Reliable Transient Applications.}
\con{In Proceedings of the IDEA League Doctoral School on Transiently Powered Computing.}
\loc{2017.}
\rem{(\chref{ch:nvmh})}
%
%
\aut{Gomez A, Conti F, Benini L.  }
\tit{Thermal Image-Based CNN's for Ultra-Low Power People Recognition.}
\con{In Proceedings of LP-EMS Workshop.}
\loc{ACM 2018.}
\rem{(\chref{ch:state_retention})}
%
\aut{Gomez A, Tretter A, Hager P, Sanmugarajah P, Benini L, Thiele L. }
\tit{Optimizing the Energy Cost of Transient Applications through Partitioning.}
\con{To be submitted.}
\loc{2018.}
\rem{(\chref{ch:state_retention})}%
%
\newpage
\noindent
The following list includes publications that were written during the PhD studies, yet are not part of this thesis.
\\%
\aut{Anagnostou P, Gomez A, Hager PA, Fatemi H , Pineda de Gyvez J, Thiele L, Benini L}
\tit{Torpor: A Power-Aware HW Scheduler for Energy Harvesting IoT SoCs. }
\con{In Proceedings of Symposium on Power and Timing Modeling, Optimization and Simulation (PATMOS). }
\loc{IEEE. 2018.}
\\%
\aut{Palossi D, Gomez A, Draskovic S, Benini L, Thiele L. }
\tit{Extending the Lifetime of Nano-Blimps via Dynamic Motor Control.}
\jou{Journal of Signal Processing Systems.}
\loc{Springer. 2018.}
\\%
\aut{Sigrist L, Gomez A, Lim R, Lippuner S, Leubin M, Thiele L.}
\tit{Measurement and validation of Energy Harvesting IoT Devices. }
\con{In Proceedings of Conference on Design, Automation \& Test in Europe (DATE).}
\loc{EDA Consortium. 2017.}
\\%
\aut{Palossi D, Gomez A, Draskovic S, Keller K, Benini L, Thiele L. }
\tit{Self-sustainability in Nano Unmanned Aerial Vehicles: A Blimp Case Study. }
\con{In Proceedings of Computing Frontiers Conference.}
\loc{ACM. 2017.}
\\%
\aut{Sigrist L, Gomez A, Lim R, Lippuner S, Leubin M, Thiele L.}
\tit{RocketLogger: Mobile Power Logger for Prototyping IoT Devices: Demo Abstract.}
\con{In Embedded Network Sensor Systems (SenSys) Conference. }
\loc{ACM. 2017.}
%
\\%
\aut{Gomez A, Bartolini A, Rossi D, Kara BC, Fatemi H, de Gyvez JP, Benini L. }
\tit{Increasing the Energy Efficiency of Microcontroller Platforms with Low-Design Margin Co-Processors. }
\jou{Microprocessors and Microsystems.}
\loc{Elsevier. 2016.}
%
\\%
\aut{Gomez A, Magno M, Lagadec MF, Benini L.}
\tit{Precise, Energy-Efficient Data Acquisition Architecture for Monitoring Radioactivity using Self-Sustainable Wireless Sensor Nodes.}
\jou{Sensors Journal.}
\loc{IEEE. 2016.}
%
\\%
\aut{Magno M, Brunelli D, Sigrist L, Andri R, Cavigelli L, Gomez A, Benini L.}
\tit{InfiniTime: Multi-sensor wearable bracelet with human body harvesting.}
\con{In Proceedings of Future Access Enablers of Ubiquitous and Intelligent Infrastructures Conference.}
\loc{Springer. 2016.}
%
\\%
\aut{Magno M, Polonelli T, Casamassima F, Gomez A, Farella E, Benini L. }
\tit{Energy-Efficient Context Aware Power Management with Asynchronous Protocol for Body Sensor Network.}
\jou{Mobile Networks and Applications.}
\loc{Springer. 2016.}
%
\\%
\aut{Sigrist L, Giannopoulou G, Huang P, Gomez A, Thiele L. }
\tit{Mixed-criticality Runtime Mechanisms and Evaluation on Multicores.}
\con{Sustainable Computing: Informatics and Systems.}
\loc{Elsevier. 2016.}
%
\\%
\aut{Gomez A, Magno M, Wen X, Benini L.}
\tit{Extending Body Sensor Nodes’ Lifetime Using a Wearable Wake-up Radio. }
\con{In Proceedings of Future Access Enablers of Ubiquitous and Intelligent Infrastructures Conference.}
\loc{Springer. 2015.}
%
\\%
\aut{Gomez A, Lagadec MF, Magno M, Benini L. }
\tit{Self-powered Wireless Sensor Nodes for Monitoring Radioactivity in Contaminated Areas using Unmanned Aerial Vehicles.}
\con{In Proceedings of Sensors Applications Symposium (SAS).}
\loc{IEEE. 2015.}
%
\\%
\aut{Gomez A, Pinto C, Bartolini A, Rossi D, Benini L, Fatemi H, de Gyvez JP.}
\tit{Reducing Energy Consumption in Microcontroller-Based Platforms with Low Design Margin Co-Processors.}
\con{In Proceedings of Conference on Design, Automation \& Test in Europe (DATE).}
\loc{EDA Consortium. 2015.}
%
\\%
\aut{Gomez A, Schor L, Kumar P, Thiele L.}
\tit{SF3P: A Framework to Explore and Prototype Hierarchical Compositions of Real-Time Schedulers.}
\con{In Proceedings of Rapid System Prototyping (RSP) Symposium.}
\loc{IEEE. 2014.}
%
\end{footnotesize}